---
title: "R Notebook"
output: html_notebook
---
Import the dataset
```{r}
setwd("C:/Users/myers/Documents/630Data")

dodgers <- read.csv("dodgers.csv")
```

```{r}
head(dodgers)
```

To start this assignment, it would be best to start looking at graphical plots to gain any possible insights.

First we will want to plot all of the main independent variables to see if there are any interesting insights that can be gained from these observations.

The first plot has to do with the time of day and attendance rates and as expected, we see a higher number of people attending the night games.
```{r}

ggplot2::ggplot(dodgers, ggplot2::aes(y=attend, x=day_night)) + 
  ggplot2::geom_point()

```

In this second plot, we want to check attendance rates and how they are related to the day of the week.

What is interesting about this plot is that it has the highest attendance rates come from Tuesday games. 
```{r}
library(ggplot2)

ggplot2::ggplot(dodgers, ggplot2::aes(x=day_of_week, y=attend)) + 
    ggplot2::geom_boxplot(alpha=0.2) +
    ggplot2::labs(x='day',
                 y='attendance numbers',
                 title='attendance by day') 
```

The third plot has to do with the opponent and attendance rates.
```{r}
ggplot2::ggplot(dodgers, ggplot2::aes(x=attend, y=opponent)) + 
        ggplot2::geom_point() + 
        ggplot2::labs(x='attendance',
                      y='opponent',
                      title = 'attendance by opponent')
```




Going forward, we will want to predict the best day to have a promo, we don't care what promo that is though. So I will just create a new field in the dataframe specifically for if a promotion occurred at all and then drop the other promotion related fields. 
```{r}
dodgers$promo <- ifelse(dodgers$cap == "YES" | dodgers$shirt == "YES" | dodgers$fireworks == "YES" | dodgers$bobblehead == "YES", "YES", "NO")

dodgers$cap <- NULL 
dodgers$shirt <- NULL
dodgers$fireworks <- NULL
dodgers$bobblehead <- NULL

head(dodgers)


```


Now that we have set up our data and drawn insights into our data, we need to create our regression model, train dataset, and test dataset, then fit them to the model.
```{r}
regr.model <- {attend ~ month + day_of_week + promo}

traintest_key <- c(rep(1, trunc((2/3)*nrow(dodgers))), rep(2, trunc((1/3)*nrow(dodgers))))

dodgers$train_test <- sample(traintest_key)

dodgers$train_test <- factor(dodgers$train_test, levels = c(1, 2), labels = c("train", "test"))
  
train <- subset(dodgers, train_test == "train")

test <- subset(dodgers, train_test == "test")

model.fit.train <- lm(regr.model, data = train)

train$attendance_predictions <- predict(model.fit.train)

test$attendance_predictions <- predict(model.fit.train, newdata = test)

accuracy <- round(cor(test$attend, test$attendance_predictions)^2, digits=3)

print(accuracy)

```
Now we will want to feed the entire dataset into the model and then we can take a look at what predictions come from this dataset.
```{r}
model.fit.all <- lm(dodger.model, data = dodgers)

print(summary(model.fit.all))
```
By looking at the above table, we can see that the biggest day of the week that has an impact on attendance is Tuesdays and in the month of June, while having a promo also has a high impact on attendance. 

My prediction would be that the best day for a promotion is a Tuesday in June based on the table above.





